---
title: "Jesse Analysis"
output: html_notebook
author: Jesse Dykes
---

# Loading necessary packages

```{r}
library(knitr)
library(tidyverse)
library(janitor)
library(here) # more easily access files in your project
library(sf) # working with simple features - geospatial
library(tmap)
library(tidycensus)
library(matrixStats)
library(sp)
```
Knitr is used for properly transforming the r notebook into an html file.
The tidyverse is used to tidy data using other packages within in such as dplyr.
The janitor packages is used to clean up the variable names from the data sets.
The here library is used to locate files on the computer to make it easier to load in the data sets.
The sf library is used to work with spatial data.
The tmap library is used to create interaactive maps.
The tidycensus data is used to get data from the U.S Census Bureau data via codes. 
The sp library allows for the use of CRS to convert degree measurements into meters.

# Load in DC Bikeshare system data and simplify the distance dataframe

```{r}
racks = st_read(here("data_raw", "Capital_Bikeshare_Locations.geojson")) |> clean_names()
racks = racks |> select(geometry,name)
```
# Load DC data

```{r}
dfShops = st_read(here("data_raw","Shopping_Centers.geojson")) |> clean_names()
pub_schools = st_read(here("data_raw", "DC_Public_Schools.geojson")) |> clean_names()
ind_schools = st_read(here("data_raw", "Independent_Schools.geojson")) |> clean_names()
```

# Simplify shops and schools dataframes

```{r}
shops = dfShops |> select(geometry,name)
p_sch = pub_schools |> select(geometry,name)
i_sch = ind_schools |> select(geometry,name)
```

# Load Api Key to access the census data

```{r}
census_api_key("9fc5d3792d3a5e922287c3f4e9995118766d50a2")
```
# Load census data

```{r}
v2018 = load_variables(2018, "acs5")
```

# Import specific census data

Vehicles per neighborhood
Population per neighborhood
Amount of those that use public transportation per neighborhood

Filter for DC and keep neighborhood shapes
Filter for years 2017-2021

```{r}

df_cencus=get_acs(geography = "tract",
                  variables=c("vehicles"="B08141_001",
                              "population" = "B01001_001",
                              "public_transportion" = "B08006_008"),
                  state="DC",geometry=TRUE,year=2021) 
```

# Clean census data

Add columns for ratios of: 
  vehicles to population
  public transport users to population
Adjust the spacial data for coordinate system 4326

```{r}
df_cens=df_cencus %>% 
  select(-moe) %>% 
  pivot_wider(names_from = "variable", 
              values_from = "estimate")

df_cens = df_cens |>
  mutate("v_to_p" = vehicles / population, "pub_transport_to_p" = public_transportion / population)

df_cens_adj = df_cens |> st_transform(4326)
```

# Turn on the interactive map from tmap

Plot the modified census data
Plot the positions of:
  bikeshare stations (transparent teal)
  shopping areas (red)
  public schools (pink)
  independent schools (violet)

```{r}

tmap_mode("view")

tm_shape(df_cens) + tm_polygons(c("v_to_p", "pub_transport_to_p"), alpha=.5) + tm_shape(shops) + tm_dots(col = "#FF1300", size = 0.1) + tm_shape(racks) + tm_dots(col = "#AEEDE2", size = 0.06, alpha = 0.5) + tm_shape(p_sch) + tm_dots(col = "#FE2DE2", size = 0.06) + tm_shape(i_sch) + tm_dots(col = "#602DE2", size = 0.06)
```

# Find the distance deltas

Use CRS to convert global degrees into meter measurements
Find the distance between each location of shops and schools and 
each bikeshare station

```{r}
crs <- CRS("+init=epsg:4326")

s_rack_d = st_distance(racks,shops)
isch_rack_d = st_distance(racks,i_sch)
psch_rack_d = st_distance(racks,p_sch)

```

# Find shop minimums

Filter the distances for the minimum value and append the corresponding shop
names to the values
Define the dataframe's type as 'shops'

```{r}
s_n = shops |> 
  st_drop_geometry() |> 
  mutate(type = "shops")
min_dist = colMins(s_rack_d)
srd_mins = cbind.data.frame(s_n,min_dist)
```

# Find public school minimums

Filter the distances for the minimum value and append the corresponding school
names to the values
Define the dataframe's type as 'public'

```{r}
p_n = p_sch |> 
  st_drop_geometry() |> 
  mutate(type = "public")
min_dist = colMins(psch_rack_d)
prd_mins = cbind.data.frame(p_n,min_dist)
```

# Find independent school minimums

Filter the distances for the minimum value and append the corresponding school
names to the values
Define the dataframe's type as 'independent'

```{r}
i_n = i_sch |> 
  st_drop_geometry() |> 
  mutate(type = "independent")
min_dist = colMins(isch_rack_d)
ird_mins = cbind.data.frame(i_n,min_dist)
```

# Combine dataframes

Combine all the dataframes into 'allmins'

Combine the school dataframes into 'schoolmins'

```{r}
allmins = rbind(srd_mins,ird_mins,prd_mins)

schoolmins = rbind(ird_mins,prd_mins)
```

# Plot

Plot the various dataframes as histograms to show the relative distances each type 
of location is from the nearest bikeshare station

Plot all and color by type

```{r}
allmins |> ggplot(aes(x=min_dist, fill = type)) + geom_histogram(binwidth = 50) + ggtitle("All locations") + xlab("Meters from Nearest Bikestation") + theme_dark()
```

Plot schools and color by type

```{r}
schoolmins |> ggplot(aes(x=min_dist, fill = type)) + geom_histogram(binwidth = 50) + ggtitle("All schools") +
  xlab("Meters from Nearest Bikestation") + theme_dark()
```

Plot the individual location dataframes and color similarly to the 'allmins' plot

```{r}
srd_mins |> ggplot(aes(x=min_dist)) + geom_histogram(binwidth = 50, color = "skyblue") + ggtitle("Shops") + ylim(0,25) +
  xlab("Meters from Nearest Bikestation") + theme_dark()

prd_mins |> ggplot(aes(x=min_dist)) + geom_histogram(binwidth = 50, color = "green") + ggtitle("Public Schools") + ylim(0,25) +
  xlab("Meters from Nearest Bikestation") + theme_dark()

ird_mins |> ggplot(aes(x=min_dist)) + geom_histogram(binwidth = 50, color = "tomato") + 
  ggtitle("Independent Schools") + xlab("Meters from Nearest Bikestation") + ylim(0,25) + 
  theme_dark()
```